\newpage
\section{Technology Primers}


You have arrived at a point at which you have completed the HDL lab and go beyond coding Verilog. Feel free to go through these technology primers if you are interested. In each tutorial you will gain an impression of what is possible to do, and why it may be useful. It will be particularly useful to you if you plan to do a master thesis in digital design at IES or generally want to continue work in this domain.

\subsection{Introduction to Gate-Level Simulation}
\begin{itemize}
\item Simulate your cpu at gate-level and verify your synthesis procedure.
\end{itemize}

\subsection{Introduction to Power Simulation}
\begin{itemize}
\item Simulate your cpu at gate-level and verify your synthesis procedure.
\end{itemize}

\subsection{Introduction to Floorplanning}
\begin{itemize}
\item Simulate after place and route
\end{itemize}

\subsection{Introduction to Constrained Random Testing}
\begin{itemize}
\item Contrained random number generation
\item Property Specification language
\end{itemize}

\subsection{Introduction to SystemVerilog Direct-Programming-Interface}
If memory bandwidth lags behind cpu performance (or you are just interested).
\begin{itemize}
\item Understand and use a high-speed Memory Controller modelled in SystemC
\end{itemize}

\subsection{Introduction to SystemC}
If memory bandwidth lags behind cpu performance (or you are just interested).
\begin{itemize}
\item Understand and use a high-speed Memory Controller modelled in SystemC
\end{itemize}

\subsection{Introduction to Matlab/Modelsim Cosimulation}
If you want to evaluate whether your cpu can meet the demands for (place your application here) during the design process (or you are just interested).
\begin{itemize}
\item Model a simple IO-module and test it (IN and OUT instructions)
\item Set up a cosimulation environment with Matlab/Modelsim
\item Record an audio stream at 8 bit PCM and increase the volume by a factor of 2 with a) Matlab and b) your cpu. Make sure to saturate the signal correctly!
\end{itemize}

\subsection{Introduction to Intellectual Property Encryption}
